Malacology: A Programmable Storage System

### 2016-10-17 paper/paper.tex
paper: apply peter's changes for services section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Mantle is a programmable load balancer that separates the metadata balancing
policies from their mechanisms. Administrators inject code to change how the
metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
authors showed how using Mantle, one can implement a single node metadata
service, a distributed metadata services with hashing, and a distributed
metadata service with dynamic subtree partitioning.

The original implementation was ``hard-coded" into Ceph and lacked robustness
(lacking things such as versioning, durability, or reliable policy
distribution).  Re-implemented using Malacology, Mantle now enjoys (1) the
versioning of MONs and (2) the durability and distribution provided by RADOS.
The version of the load balancer, which is stored in the MDS map, corresponds
to an object name in RADOS that houses the load balancer code.

@@

Ensuring that the version of the current load balancer is consistent across the
MDS cluster was ignored in the original implementation. The user had to set the
version on each individual MDS and it was trivial to make the version
inconsistent across MDS nodes.

With Malacology, Mantle stores the version of the current load balancer in the
MDS map.  Similar to the PG map or MON maps, the MDS map has all the
information about the topology and state of the MDS cluster.  Recall that the
MONs propogate these maps across the cluster and ensure that the map versions
are consistent with PAXOS.  While storing the code snippets that define the
policies in the MDS map is possible, it is not recommended as it is an
expensive operation (a PAXOS voting round).

@@

\subsection{ZLog: A Fast Distributed Shared-Log}
\label{sec:zlog}

The second service implemented on Malacology is ZLog, a high-performance
distributed shared-log that is based on the CORFU
protocol~\cite{balakrishnan_corfu_2012}. A
shared-log is a general, yet powerful abstraction used to build many
distributed systems. However, existing implementations that rely on consensus
algorithms such as PAXOS funnel I/O through a single point introducing a
bottleneck that restricts throughput. In contrast, the CORFU protocol is able
to achieve high throughput using a network counter service, called a 
\emph{sequencer}, that avoids all persistent I/O in the common case.

While a full description of the CORFU system is beyond the scope of this
paper, we'll briefly describe the custom storage device interface, sequencer
service, and recovery protocol, and how these services are instantiated in the
Malacology framework.

@@

High-performance in CORFU is achieved using a sequencer service that assigns
log positions to clients by reading from a volatile, in-memory counter which
can run at a very high throughput and at low-latency. And since the sequencer
is centralized, ensuring serializability in the common case is trivial.  The
primary challenge in CORFU is handling the failure of the sequencer in a way
that preserves correctness. Failure of the sequencer service in CORFU is
handled by a recovery algorithm that recomputes the new sequencer state using
an application-specific custom storage interface to discover the tail of the
log, while simultaneously invalidating stale client requests using an
epoch-based protocol.

@@

In Section~\ref{evaluation} we quantify the trade-offs of throughput and
latency for an approach based on a round-robin batching mode, and compare this
mode to one in which the metadata server mediates access to the sequencer
state when in it is being shared among multiple clients.

@@

In order to repopulate the cached tail value during recovery of a sequencer,
the maximum position in the log must be obtained. To do this, the storage
interface exposes an additional \emph{seal} method that atomically installs a
new epoch value and returns the maximum log position that has been written.
Since the recovery protocol does not inform clients of a new epoch value until
the end of the process, the validity of the tail can be guaranteed despite the
iterative process of contacting storage devices.  Recovery of the sequencer
process itself may be handled in many ways, such as leader election using an
auxiliary service such as PAXOS.

\section{Evaluation}
\label{evaluation}

@@

\subsection{Mapping ZLog onto Malacology}
\label{mapping-zlog-onto-malacology}
We evaluate the mapping of ZLog onto Malacology by examining the costs of
utilizing the metadata service as a sequencer, and propagating interface
changes throughout the cluster using the monitoring service. We have omitted
evaluation of the overhead costs of adding object interfaces themselves due to
lack of space, and the fact that they are heavily tested and used already in
production with acceptable performance, introducing a small constant overhead
per operation but otherwise does not affect scalability.

>>>>

Mantle is a programmable load balancer that separates the metadata balancing
policies from their mechanisms. Administrators inject code to change how the
metadata cluster distributes metadata. In ~\cite{sevilla:sc15-mantle} the
authors showed how using Mantle, one can implement a single node metadata
service, a distributed metadata service with hashing, and a distributed
metadata service with dynamic subtree partitioning.

The original implementation was ``hard-coded" into Ceph and lacked robustness
(no versioning, durability, or policy distribution).  Re-implemented using
Malacology, Mantle now enjoys (1) the versioning provided by the monitor
daemons and (2) the durability and distribution provided by Ceph's reliable
object store.

@@

Ensuring that the version of the current load balancer is consistent across the
physical servers in the metadata cluster was ignored in the original
implementation. The user had to set the version on each individual server and
it was trivial to make the version inconsistent.

With Malacology, Mantle stores the version of the current load balancer in the
Ceph map using the service metadata interface. The version of the load balancer
corresponds to an object name in the balancing policy. Using the service
metadata interface ensures that all metadata servers use the same version of
the load balancer. As a result, the policy version inherits the consensus and
consistency of Ceph's internal monitor daemons.

@@

\subsection{ZLog: A Fast Distributed Shared Log}
\label{sec:zlog}

The second service implemented on Malacology is ZLog, a high-performance
distributed shared-log that is based on the CORFU
protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
powerful abstraction used to build many distributed systems, such as
cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
storage engines~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
However, existing implementations that rely on consensus algorithms such as
PAXOS funnel I/O through a single point introducing a bottleneck that restricts
throughput.  In contrast, the CORFU protocol is able to achieve high throughput
using a network counter service, called a \emph{sequencer}, for obtaining the
entry at the tail of log. This metadata operation that avoids all persistent
I/O in the common case.

While a full description of the CORFU system is beyond the scope of this
paper, we briefly describe the custom storage device interface, sequencer
service, and recovery protocol, and how these services are instantiated in the
Malacology framework.

@@

High-performance in CORFU is achieved using a sequencer service that assigns
log positions to clients by reading from a volatile, in-memory counter which
can run at a very high throughput and at low latency. Since the sequencer
is centralized, ensuring serializability in the common case is trivial.  The
primary challenge in CORFU is handling the failure of the sequencer in a way
that preserves correctness. Failure of the sequencer service in CORFU is
handled by a recovery algorithm that recomputes the new sequencer state using
an application-specific custom storage interface to discover the tail of the
log, while simultaneously invalidating stale client requests using an
epoch-based protocol.

@@

In Section~\ref{evaluation} we quantify the trade-offs of throughput and
latency for an approach based on a round-robin batching mode, and compare this
mode to one in which the metadata server mediates access to the sequencer state
when in it is being shared among multiple clients. Quantifying these trade-offs
should provide administrators with guidelines for setting the tunables for
different ``caching" modes of the sequencer.

@@

In order to repopulate the cached tail value during recovery of a sequencer,
the maximum position in the log must be obtained. To do this, the storage
interface exposes an additional \emph{seal} method that atomically installs a
new epoch value and returns the maximum log position that has been written.
Since the recovery protocol does not inform clients of a new epoch value until
the end of the process, the validity of the tail can be guaranteed despite the
iterative process of contacting storage devices.  Recovery of the sequencer
process itself may be handled in many ways, such as leader election using an
auxiliary service such as PAXOS. In our implementation, the recovery is the
same (and is inherited from) the Ceph metadata service.

\section{Evaluation}
\label{sec:evaluation}

@@

\subsection{Mapping ZLog onto Malacology}
\label{sec:mapping-zlog-onto-malacology}
We evaluate the mapping of ZLog onto Malacology by examining the costs of
utilizing the metadata service as a sequencer, and propagating interface
changes throughout the cluster using the monitoring service. We have omitted
evaluation of the overhead costs of adding object interfaces themselves due to
lack of space, and the fact that they are heavily tested and used already in
production with acceptable performance, introducing a small constant overhead
per operation but otherwise does not affect scalability.
